---
title: "HW3-Darshan Patel-3:30-5:30PM"
author: "Darshan Patel"
date: "10/18/2018"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In the beginning of the 17th century, John Graunt wanted to determine the effect of the plague ont he population of England; two hundred years later, Pierre-Simon Laplace wanted to estimate the population of France. Both Graunt and Laplace implemented what is now called the `capture-recapture methood.` This technique is used to not only count human populations (such as the homeless) but also animals in the wild. 

In its simplest form, $n_1$ individuals are "captured," "tagged,"" and released. A while later, $n_2$ individuals are "captured" and the number of "tagged" individuals, $m_2$ is counted. If $N$ is the true total population size, we can estimate it with $\hat{N}_{LP}$ as follows: $$ \hat{N}_{LP} = \frac{n_1 n_2}{m_2} $$ 
using the relation $$ \frac{n_1}{N} = \frac{m_2}{n_2} $$ This is called the Lincoln-Peterson estimator. 

We make several strong assumptions when we use this method: (a) each individual is independently captured, (b) each individual is equally likely to be captured, (c) there are no births, deaths, immigration, or emigration of individuals (i.e., a closed population), and (d) the tags do not wear off (if it is a physical mark) and no tag goes unnoticed by a researcher. 

_Goal:_ In this assignment, you will develop a Monte-Carlo simulation of the capture-recapture method and investigate the statistical properties of the Lincoln-Peterson and Chapman estimators of population size $N$. (Since you are simulating your own data, you know the true value of the population size $N$ allowing you to study how well these estimators work.)

```{r}
# Import libraries
library(tidyverse)
```

## Question 1: 
Simulate the capture-recapture method for a population of size $N = 5,000$ when $n_1 = 100$ and $n_2 = 100$ using the `sample()` function (assume that each individual is equally likely to be "captured"). Determine $m_2$ and calculate $\hat{N}_{LP}$. 

Answer: 
```{r}
# Create variables
n <- 5000
n_1 <- 100
n_2 <- 100

# Sample the captured points
cap_one <- sample(1:n, n_1, replace = FALSE)
cap_two <- sample(1:n, n_2, replace = FALSE)

# Calculate m_2 and N_LP
m_2 <- length(intersect(cap_one, cap_two))
N_LP <- n_1 * n_2 / m_2
paste("m_2:", m_2, sep = ' ')
paste("N_LP:", N_LP, sep = ' ')
```

## Question 2: 
Write a function to simulate the capture-recapture procedure using the inputs: $N$, $n_1$. $n_2$ and the number of simulation runs. The function should output in list form (a) a data frame with two columns: the values of $m_2$ and $\hat{N}_{LP}$ for each iteration and (b) $N$. Run your simulation for $1,000$ iterations for a population of size $N = 5,000$ where $n_1 = n_2 = 100$ and make a histogram of the resulting $\hat{N}_{LP}$ vector. Indicate $N$ on the plot. 

Answer: 
```{r}
# Function to simulate the capture-recapture procedure using the inputs: N, n_1, n_2 and number of simulation runs
capture_recapture = function(N, n_1, n_2, runs){
  
  # Create vector to store m_2s and N_LPs
  m_2 <- c()
  N_LP <- c()
  
  # For each run, perform the capture-recapture procedure 
  for(i in 1:runs){
    
    # Sample the captured points 
    cap_one <- sample(1:N, n_1, replace = FALSE)
    cap_two <- sample(1:N, n_2, replace = FALSE)
    
    # Finds the points that are recaptured 
    recaptured <- length(intersect(cap_one, cap_two))
    
    # Append to m_2 vector
    m_2 <- c(m_2, recaptured)
    
    # Calculate N_LP and append to its vector 
    N_LP <- c(N_LP, (n_1 * n_2 / recaptured))
  }
  
  # Return a list where the first element is a dataframe of m_2 and N_LP 
  # and the second element is the actual N
  return(list(df = data.frame(m_2, N_LP), N = N))
}
```

Simulation:
```{r}
# Perform a simulation using 1000 runs
simulation <- capture_recapture(5000, 100, 100, 1000)
df <- simulation$df

# Plot a histogram of the calculated N_LP values
ggplot(df, aes(x = df$N_LP)) + geom_histogram(color = "darkblue", fill = "lightblue") + ggtitle("Calculated N_LP value from \n Simulation of Capture-Recapture Procedure") + geom_vline(aes(xintercept = 5000), color = "black", linetype = "dashed") + geom_text(aes(x = simulation$N - 800, y = 280, label = "actual N")) + labs(x = "N_LP")
```


## Question 3: 
What percent of the estimated population values in Question 2 were infinite? Why can this occur? 

Answer: 
```{r}
per_inf <- paste(100 * length(which(is.infinite(simulation$df$N_LP))) / nrow(simulation$df), "%", sep = ' ')
per_inf
```
The percentage of infinite $N_{\text{LP}}$ calculated is `r per_inf`. This occurs because in some rare cases, there was no recapture of any points, meaning $m_2 = 0$. Thus in the calculation of $N_{LP}$, there is a $0$ in the denominator, causing R to give an infinite result.


## Question 4:
An alternative to the Lincoln-Peterson estimator is the Chapman estimator:
$$ \hat{N}_C = \frac{(n_1 + 1)(n_2 + 1)}{m_2 + 1} - 1$$ 
Use the saved $m_2$ values from question 2 to compute the corresponding Chapman estimates for each iteration of your simulation. Construct a histogram of the resulting $\hat{N}_C$ estimates, indicating $N$ on the plot. 

Answer: 
```{r}
# Calculate N_C from the m_2 values from the previous simulation runs 
N_C <- data.frame(chapman = (((n_1 + 1)*(n_2 + 1))/(simulation$df$m_2 + 1)) - 1)

# Plot a histogram of the calculated N_C values
ggplot(N_C, aes(x = N_C$chapman)) + geom_histogram(color = "steelblue4", fill = "skyblue") + ggtitle("Calculated N_C from \n Simulation of Capture-Recapture Procedure") + geom_vline(aes(xintercept = 5000), color = "black", linetype = "dashed") + geom_text(aes(x = simulation$N - 800, y = 300, label = "actual N")) + labs(x = "N_C")
```

## Question 5: 
An estimator is considered _unbiased_ if, on average, the estimator equals the true population value. For example, the sample mean $\bar{x} = \sum_{i=1}^n x_i / n$ is unbiased because on average the sample mean $\bar{x}$ equals the population mean $\mu$ (i.e., the sampling distribution is centered around $\mu$). This is a desirable property for an estimator to have because it means our estimator is not systematically wrong. To show that an estimator $\hat{\theta}$ is an unbiased estimator of the true value $\theta$, we would need to mathematically prove that $\mathrm{E}[\hat{\theta}] - \theta = 0$ where $\mathrm{E}[\cdot]$ is the expectation (i.e., theoretical average). Instead, we will investigate this property empirically by replacing the theoretical average $\mathrm{E}[\hat{\theta}]$ with the sample average of the $\hat{\theta}$ values from our simulation (i.e., $\sum_{i=1}^{n_{\text{sim}}} \hat{\theta} / n_{\text{sim}}$ where $n_{\text{sim}}$ is the number of simulation runs; $\theta$ is $N$ in this case and $\hat{\theta}$ is either $\hat{N}_{LP}$ or $\hat{N}_C$ as both are ways to estimate $N$).

Estimate the bias of the Lincoln-Peterson and Chapman estimators based on the results of your simulation. Is either estimator unbiased when $n_1, n_2 = 100$? 

Answer: Note that to compute the mean of $N_LP$, the infinites should be ignored and $n_{\text{sim}}$ should be properly instated.  
```{r}
# Isolate the finite N_LP values
N_LP_finite <- simulation$df$N_LP[is.finite(simulation$df$N_LP)]

# Calculate the bias of N_LP
bias_LP <- mean(N_LP_finite) / length(N_LP_finite)

# Calculate the bias of N_C
bias_chapman <- mean(N_C[,1]) / length(N_C[,1])
paste("The bias of the Lincoln-Peterson estimator is", bias_LP, sep = ' ')
paste("The bias of the Chapman estimator is", bias_chapman, sep = ' ')
```
Neither of the estimators are unbiased when $n_1, n_2 = 100$. 

## Question 6:
Based on your findings, is the Lincoln-Peterson or Chapman estimator better? Explain your answer. 

Answer: Based on the findings, the Chapman estimator does better to estimate $N$ than the Lincoln-Peterson estimator. One reason is that the bias of the estimator calculated was lower than the one calculated using the Lincoln-Peterson estimator. Another reason is that the Lincoln-Peterson estimator was accustomed to calculating infinities which is theoretically not possible in real life scenarios. 


## Question 7:
Explain why the assumptions (a), (b) and (c) used to make the Lincoln-Peterson estimator are unrealisitic. 

Answer: Unrealistic Assumptions: 

* A: When one individual is captured, it is possible that another individual is captured in a nearby vicinity as well 
* B: Each individual may have factors that influence its chances of being captured, such as health or agility
* C: Births and deaths events cannot be held constant in the real world as they are natural events.

